---
title: "Covid91 vaccine study Final2021"
author: "Jocelyn Lazcano"
date: "Spring 2021"
output: 
  html_document:
    number_sections: true
    toc: true

---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="covid91.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups consisting of Males, Females, just the drug users, and just those in the LGBTQ community. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Methods 

The first step step is isolating subpopulations

```{r}
Maledata <- filter(FinalData,sex=="guy")
Maledata$sex <- as.factor(Maledata$sex)
Maledata$infected=factor(Maledata$infected)
Maledata$infected<-relevel(Maledata$infected, "OK")
```


# Males

## Barchart

```{r}
males <- maledata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))

basicC <- ggplot(males,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")
#Now for percentage plot

basicCC <- ggplot(males,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This bar chart appears to show that around 340 people who were infected with Covid-19, 57% of males who were given the drug. While 43% that were infected were given the placebo. 

### Numerical Summaries  

```{r}
table1 <- xtabs(~treatment + infected, data=maledata)
rowPerc(table1)
colPerc(table1)
print(table1)
```

The table shows that about 340 males were infected with Covid-19 out of approximately 10,400 males in the study. The table shows that 3.72% of males who recieved the drug became infected with Covid.While  only 2.55% of the males who recived the placebo contracted Covid. The table also shows that males that were vaccinated were 50% likely to contract Covid. This shows that the vaccine is not effective in males. 


## Inferential  

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The Chi Squared table shows that the P-Value of 0.0006181 is so small that it is highly unlikely that the results are by choice.
$H_0$ the null hypothesis given taking the vaccine there is no difference of whether males can contract Covid or not. 
$H_A$ the alternative hypothesis is that the vaccine in fact does not make a difference in whether or not a male contracts covid-19. Therefore we reject the null hypothesis.

The odds ratio from the Fisher Exact test is 1.471:1. When taking the odds ratio and computing it into a percent of effectiveness by using the formula 0/(1+0)*100 you get 59.5% effectiveness. 


Ultimately the vaccine is 40% effective in the male population. Which means it is ineefective.


# Females

```{r}
Femaledata <- filter(FinalData,sex=="gal")
Femaledata$sex <- as.factor(Femaledata$sex)
Femaledata$infected=factor(Femaledata$infected)
Femaledata$infected<-relevel(Femaledata$infected, "OK")
levels(Femaledata$infected)
```


## Barchart

```{r}
Females <- Femaledata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))

basicC <- ggplot(Females,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")
#Now for percentage plot

basicCC <- ggplot(Females,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

The barchart shows that out of approximaetly 607 females who were infected with COVID-19, only 32% were given the drug and the other 66% were given the placebo. 


### Numerical Summaries  

```{r}
table2 <- xtabs(~infected + treatment, data=Femaledata)
rowPerc(table2)
colPerc(table2)
print(table2)
```

The table shows that 607 females out of 9,200 were infected with COVID-19. The table also shows that 4.44% of the women who received the drug contracted COVID-19. While the 8.67% of women who got the placebo contracted COVID-19. Women who received this vaccine were 67% less likely to contract COVID-19. This means that the vaccine is effective in females. 

## Inferential  

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```

The p-value is so small at 0.000000000000002 that it is highly unlikely that the results occurred by chance.

$H_U$ the alternative hypothesis is that the vaccine in fact does make a difference in whether or not females contract COVID_19.

Therefore, we reject the null hypothesis.

The Fisher Exact test odds ration is 2.05:1 when taking the odds ratio and computing it into a percent of effectiveness by using the formula 0/(1+0)*100 the result is 67% effectiveness. 

The vaccine is 67% effective in females. 

# LGBTQ

## Barchart

```{r}
males <- maledata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))

basicC <- ggplot(males,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")
#Now for percentage plot

basicCC <- ggplot(males,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```


# Drug Users

```{r}
Druggiedata <- filter(FinalData,DrugUser=="yes")
Druggiedata$sex <- as.factor(Druggiedata$sex)
```


## Barchart

```{r}
Druggies <- Druggiedata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))

basicC <- ggplot(Druggies,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")
#Now for percentage plot

basicCC <- ggplot(Druggies,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

### Numerical Summaries  

```{r}
table3 <- xtabs(~infected + treatment, data=Femaledata)
rowPerc(table3)
colPerc(table3)
print(table3)
```

## Inferential  

```{r}
chisq.test(table3)
chisqtestGC(table3)
fisher.test(table3)
```

The vaccine is 58% effective in the Druggie population. 





# Discussion & Conclusion
