---
title: "Covid91 vaccine study Final2021"
author: "Jocelyn Lazcano"
date: "Spring 2021"
output: 
  html_document:
    number_sections: true
    toc: true

---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="covid91.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups consisting of Males, Females, just the drug users, and just those in the LGBTQ community. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Methods 

The first step step is isolating subpopulations

```{r}
Maledata <- filter(FinalData,sex=="guy")
Maledata$sex <- as.factor(Maledata$sex)
Maledata$infected=factor(Maledata$infected)
Maledata$infected<-relevel(Maledata$infected, "OK")
```


# Males

## Barchart

```{r}
Males <- Maledata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))

basicC <- ggplot(Males,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(Males,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This bar chart appears to show that around 340 people who were infected with Covid-19, 57% of males who were given the drug. While 43% that were infected were given the placebo. 

### Numerical Summaries  

```{r}
table1 <- xtabs(~treatment + infected, data=Maledata)
rowPerc(table1)
colPerc(table1)
print(table1)
```

The table shows that about 340 males were infected with Covid-19 out of approximately 10,400 males in the study. The table shows that 3.72% of males who recieved the drug became infected with Covid.While  only 2.55% of the males who recived the placebo contracted Covid. The table also shows that males that were vaccinated were 50% likely to contract Covid. This shows that the vaccine is not effective in males. 


## Inferential  

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The Chi Squared table shows that the P-Value of 0.0006181 is so small that it is highly unlikely that the results are by choice.
$H_0$ the null hypothesis given taking the vaccine there is no difference of whether males can contract Covid or not. 
$H_A$ the alternative hypothesis is that the vaccine in fact does not make a difference in whether or not a male contracts covid-19. Therefore we reject the null hypothesis.

The odds ratio from the Fisher Exact test is 1.471:1. When taking the odds ratio and computing it into a percent of effectiveness by using the formula 0/(1+0)*100 you get 59.5% effectiveness. 


Ultimately the vaccine is 40% effective in the male population. Which means it is ineefective.


# Females

```{r}
Femaledata <- filter(FinalData,sex=="gal")
Femaledata$sex <- as.factor(Femaledata$sex)
Femaledata$infected=factor(Femaledata$infected)
Femaledata$infected<-relevel(Femaledata$infected, "OK")
levels(Femaledata$infected)
```


## Barchart

```{r}
Females <- Femaledata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))

basicC <- ggplot(Females,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")
#Now for percentage plot

basicCC <- ggplot(Females,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

The barchart shows that out of approximaetly 607 females who were infected with COVID-19, only 32% were given the drug and the other 66% were given the placebo. 


### Numerical Summaries  

```{r}
table2 <- xtabs(~infected + treatment, data=Femaledata)
rowPerc(table2)
colPerc(table2)
print(table2)
```

The table shows that 607 females out of 9,200 were infected with COVID-19. The table also shows that 4.44% of the women who received the drug contracted COVID-19. While the 8.67% of women who got the placebo contracted COVID-19. Women who received this vaccine were 67% less likely to contract COVID-19. This means that the vaccine is effective in females. 

## Inferential  

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```

The p-value is so small at 0.000000000000002 that it is highly unlikely that the results occurred by chance.

$H_U$ the alternative hypothesis is that the vaccine in fact does make a difference in whether or not females contract COVID_19.

Therefore, we reject the null hypothesis.

The Fisher Exact test odds ration is 2.05:1 when taking the odds ratio and computing it into a percent of effectiveness by using the formula 0/(1+0)*100 the result is 67% effectiveness. 

The vaccine is 67% effective in females. 

# LGBTQ

```{r}
LGBTQdata <- filter(FinalData, LGBTQ== "gay")
LGBTQdata$DrugUser <- as.factor(LGBTQdata$LGBTQ)
LGBTQdata$infected=factor(LGBTQdata$infected)
LGBTQdata$infected<-relevel(LGBTQdata$infected, "OK")
```


## Barchart

```{r}
LGBTQ <- LGBTQdata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))

basicC <- ggplot(LGBTQ,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")
#Now for percentage plot

basicCC <- ggplot(LGBTQ,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This bar chart shows that shows that out of the approximately 116 LGBTQ, who were infected with COVID-19. 70% were given the drug, and 29% were given the placebo. 


### Numerical Summaries  

```{r}
table3 <- xtabs(~infected + treatment, data=LGBTQdata)
rowPerc(table3)
colPerc(table3)
print(table3)
```

The table shows that about 116 LGBTQ were infected with COVID-19 out of approximately 2,040 LGBTQ in the study. The table also shows that 70% of LGBTQ who received the drug contracted COVID-19 and 29% of LGBTQ who received the placebo contracted COVID-19. Lastly, the table show that LGBTQ who got the vaccine were about 29% more likely to contract COVID. This means the placebo is not effective for the LQBTQ population. 


## Inferential  

```{r}
table3 <- xtabs(~treatment + infected, data= LGBTQdata)
chisq.test(table3)
chisqtestGC(table3)
fisher.test(table3)
```

This table shows that about 116 LGBTQ had been infected with the COVID virus out of 2,040 participants in this study. This table also shows that 70% of LGBTQ who received the drug contracted COVID and 29% who received the placebo contracted COVID. Finally, the table shows that LGBTQ who got their vaccine were 29% more likely to contract COVID_19. This shows that the vaccine is ineffective for the LGBTQ popultion. 



# Drug Users

```{r}
Druggiedata <- filter(FinalData,DrugUser=="yes")
Druggiedata$sex <- as.factor(Druggiedata$sex)
```


## Barchart

```{r}
Druggies <- Druggiedata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))

basicC <- ggplot(Druggies,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")
#Now for percentage plot

basicCC <- ggplot(Druggies,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This barchart shows that out of the approximately 117 DrugUsers who contracted COVID-19, 32% were given the drug. While 66% were given the placebo.

### Numerical Summaries  

```{r}
table4 <- xtabs(~infected + treatment, data=Druggiedata)
rowPerc(table4)
colPerc(table4)
print(table4)
```

The table shows that 117 of drug users out of 2,280 were infected with COVID-19. The table shows that 33.3% of drug users who received the drug contracted Covid-19 and 66.6% of drug users who received the placebo contracted covid. Drug users who got the vaccine the vaccine were 60% less likely to get COVID-19. Thus making the vaccine effetive for drug users. 


## Inferential  

```{r}
chisq.test(table4)
chisqtestGC(table4)
fisher.test(table4)
```

The P-value is so small at 0.002273 that it is highly unlikely that the results were by chance. 

$H_0$ the null hypothesis is that the vaccine makes no difference in the drug user population when it comes to contracting COVID-19.

$H_0$ the alternative hypothesis is that the vaccine in fact does make a difference in whether or not a drug user contracts COVID-19.

We reject the null hypothesis.

The fisher exact test odds ratio is 1.831:1 when taking the odds ratio and computing it into a percent of effectiveness by using the formula 0/(1+0)*100 you get 60% effectiveness. Utimately, the vaccine is 60% effective in the Druggie population.



# Discussion & Conclusion

When comparing the four populations, we find that two of populations had the highest rate of effectiveness were the Female group at 67% and the DrugUser group at 60% The Male and LGBTQ groups showed that the vaccine was ineffective. The Male group had 40% effectiveness while the LGBTQ group had 29% effectiveness. 
